On the adhesivity of EGGs

### 2025-03-03 calco/main.tex
mi fermo qua per oggi

\begin{restatable}{lemma}{kpp}\label{lemma:kern_pairs_pres_pullbacks}
	Let $f\colon X \to Y$ and $g\colon Z \to W$ be two arrows admitting kernel pairs and suppose that the solid part of the three squares below is given. 
	%
	If the leftmost square is commutative, then there exists a a unique arrow $k_h: K_f \to K_g$ making the other two commutative.
	%
	Moreover, if the leftmost is a pullback, then also the other two are so.
%
	\[\xymatrix@R=15pt{X \ar[r]^{h}\ar[d]_{f}& Z \ar[d]^{g} & K_f \ar@{.>}[r]^{k_h} \ar[d]_{\pi^1_f}& K_g \ar[d]^{\pi^1_g} & K_f  \ar@{.>}[r]^{k_h} \ar[d]_{\pi^2_f}& K_g \ar[d]^{\pi^2_g}\\ Y \ar[r]_t& W & X \ar[r]_{h} & Z & X\ar[r]_{h} &Z}\]

\end{restatable}
%[Proof in \Cref{kpp-proof}]

The previous result allows us to deduce the following lemma in an $\mathcal{M}$-adhesive context.

\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

@@

\begin{restatable}{corollary}{mn}\label{cor:mono2}
	An arrow $(h_E, h_V, h_Q): \mathcal{G\to H}$ in $\EqHyp$ is a regular mono if and only if $h_E, h_V$ and $h_C$ are all monos.
\end{restatable}

@@

>>>>

\begin{restatable}{lemma}{kpp}\label{lemma:kern_pairs_pres_pullbacks}
	Let $f\colon X \to Y$ and $g\colon Z \to W$ be two arrows admitting kernel pairs and suppose that the solid part of the three squares below is given.

	If the leftmost square is commutative, then there exists a a unique arrow $k_h: K_f \to K_g$ making the other two commutative.
	\[\xymatrix@R=15pt{X \ar[r]^{h}\ar[d]_{f}& Z \ar[d]^{g} & K_f \ar@{.>}[r]^{k_h} \ar[d]_{\pi^1_f}& K_g \ar[d]^{\pi^1_g} & K_f  \ar@{.>}[r]^{k_h} \ar[d]_{\pi^2_f}& K_g \ar[d]^{\pi^2_g}\\ Y \ar[r]_t& W & X \ar[r]_{h} & Z & X\ar[r]_{h} &Z}\]

		Moreover, the following hold true:
		\begin{enumerate}
			\item if the leftmost squareis a pullback, then also the other two are so;
			\item if $h$ is mono then $k_h$ is mono too.
		\end{enumerate} 
\end{restatable}
%[Proof in \Cref{kpp-proof}]

The previous result allows us to deduce the following lemma in an $\mathcal{M}$-adhesive context.
\todo{A questo punto la proposizione seguente possiamo anche buttarla via se serve recuperare spazio}
\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\todo{Inserire bla bla}

\noindent
\parbox{7.5cm}{
	\begin{restatable}{proposition}{mps}\label{prop:kerset}
		Suppose that in $\Set$ the cube aside is given, whose top and bootom faces are pushout and such that $m\colon A\mto C$ and $m'\colon A'\mto C'4$ are injections. Then the right square, made by the kernel pairs of the vertical arrows is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\begin{restatable}{corollary}{mn}\label{cor:mono2}
	If $(h_E, h_V, h_Q): \mathcal{G\to H}$ is a regular mono in $\EqHyp$, then $h_E, h_V$ and $h_C$ are all monos.
\end{restatable}

@@

To prove \Cref{prop:kerset} we need some results about pushouts and coproducts in $\Set$.

\begin{lemma}\label{lem:po_set}
	The following are true:
	\begin{enumerate}
		\item if $(P, \{p_i\}_{i\in I})$ is a coproduct, then for every arrow $f\colon 1\to P$ there is exactly one $k\in I$ and $f_k\colon 1\to P_k$ such that $f=p_k\circ f_k$;

		\parbox{11cm}{
		\item suppose that the square on the right is a pushout and let $\iota_{m}\colon E\mto B$ be the inclusion of the complement of the image of $m$, then $(D, \{n, h\circ \iota_m \})$ is a coproduct; in particular, $h\circ \iota_f$ is mono.}\hfill \parbox{2cm}{\xymatrix@R=15pt{A \ar@{>->}[r]^{m}  \ar[d]_{g}& B \ar[d]^{h}\\ C \ar@{>->}[r]_{n} & D}}
	\end{enumerate}
\end{lemma}
\begin{proof}
	\begin{enumerate}
		\item This is obvious.
		\item Let $f\colon E\to Z$ and $k\colon C\to Z$ be two arrows with the same codomain. . By definition of complement and since $m$ is mono we know that $(B, \{m, \iota_m\})$ is a coproduct. Define $\phi\colon B\to Z$ as the unique arrow such that \[\phi\circ m=k\circ g \qquad \phi \circ \iota_m=f\] 
		By construction there is a unique arrow $\psi \circ D\to Z$ such that 
		\[\psi \circ n=k \qquad \psi \circ h=\phi \]
		But this entails that 
		\[
		\psi \circ h\circ \iota_m =\phi \circ \iota_m=f\]

		On the other hand, if $\psi'\colon D\to Z$ is another arrow such that
		\[\psi' \circ n=k \qquad \psi' \circ h\circ \iota_m=f \]
		then we have
		\[\psi'\circ h\circ m=\psi'\circ n\circ g=k\circ g=\phi \circ m =\psi \circ h\circ n \]
		showing $\psi'=\psi$. \qedhere
	\end{enumerate}
\end{proof}

\noindent
\parbox{7.5cm}{
\mps*}
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

\begin{proof} 
	Let $u\colon K_b\to Z$ and $v\colon K_c\to Z$ be two arrows such that $u\circ k_{f'}=v\circ k_{m'}$. We want to construct an arrow $w\colon K_d\to Z$ such that 
	\[w\circ k_{g'}=v \qquad w\circ k_{n'}=u\]

	Now, if $E'\subseteq C'$ is the complement of the image of $m'$ and $\iota\colon E'\mto B'$ the corresponding inclusion, then by \Cref{lem:po_set} we know that $(D', \{n', g'\circ \iota\})$ is a coproduct. Thus if $x$ is an element of $K_d$, we have exactly four cases.

	\begin{enumerate}
		\item $\pi^1_d(x)$ and $\pi^2_d(x)$ both belong to the image of $n'$. Let $x_1$ and $x_2$ be such that $\pi^1_d(x)=n'(x_1)$ and $\pi^2_d(x)=n'(x_2)$, then 
		\[n(b(x_1))=d(n'(x_1))=d(\pi^1_d(x))=d(\pi^1_d(x))=d(n'(x_2))=n(b(x_2)) \]
		Since $n$ is mono we deduce that $b(x_1)=b(x_2)$ and so there exists $y\in K_{b}$ such that $\pi^1_d(y)=x_1$ and $\pi^2_d(x)=x_2$. But then we have
		\begin{align*}
	\pi^1_d(k_{n'}(y))&=n'(\pi^1_d(y)) =n'(x_1)=\pi^1_d(x)\\
	\pi^2_d(k_{n'}(y))&=n'(\pi^2_d(y)) =n'(x_2)=\pi^2_d(x)
		\end{align*}
		Thus, by \Cref{lemma:kern_pairs_pres_pullbacks} $y$ is the unique element $k_{n'}(y)=x$. We can thus define $w(x)=u(y)$. 
		\item $\pi^1_d(x)$ and $\pi^2_d(x)$ both belong to the image of $g'\circ \iota$. Consider $x_1$ and $x_2$ in $E'$ such that
		\[g'(\iota(x_1))=\pi^1_d(x) \qquad  g'(\iota(x_2))=\pi^1_d(x)\]
		Thus we have
		\begin{align*}
			g(c(\iota(x_1)))=d(g'(\iota(x_1)))=d(\pi^1_d(x))=d(\pi^2(x))=
		\end{align*}

		\item $\pi^1_d(x)$ belongs to the image of $n'$  and $\pi^2_d(x)$ to that of $g'\circ \iota$;
		\item $\pi^2_d(x)$ belongs to the image of $n'$  and $\pi^1_d(x)$ to that of $g'\circ \iota$;
	\end{enumerate}

	To prove uniqueness, let we split the cases as above.
	\begin{enumerate}
		\item 
		\item 
		\item 
		\item \qedhere
	\end{enumerate}

	Since $\Set$ is cocomplete, we can take a pushout square as the one  on the right below.

	\parbox{10.5cm}{Now, notice that \begin{align*}
		n'\circ \pi^1_{b}\circ k_{f'}&=n'\circ f'\circ \pi^1_{a}=g'\circ m'\circ \pi^1_{a}=g'\circ \pi^{1}_{c}\circ k_{m'}\\
		n'\circ \pi^2_{b}\circ k_{f'}&=n'\circ f'\circ \pi^2_{a}=g'\circ m'\circ \pi^2_{a}=g'\circ \pi^{2}_{c}\circ k_{m'}
		\end{align*}}\hfill \parbox{2cm}{\xymatrix@R=15pt{K_a \ar[r]_{k_{f'}} \ar[d]_{k_{m'}}& K_b  \ar[d]^{p_2}\\ K_c \ar[r]^{p_1} & P}}

	In particular, this entails that there are two arrows $\pi_1, \pi_2\colon P\rightrightarrows D'$ such that 
	\[\pi_1\circ p_1=g'\circ \pi^1_c \quad \pi_1\circ p_2=n'\circ \pi^1_b \quad \pi_2\circ p_1=g'\circ \pi^2_c \quad \pi_2\circ p_2=n'\circ \pi^2_b\]

 Moreover,  we have
 \begin{align*}
 	d\circ \pi_1\circ p_1&=d\circ g'\circ \pi^1_c=g\circ c\circ \circ \pi^1_c=g\circ c\circ \pi^2_c=d\circ g'\circ \pi^2_c=d\circ \pi_2 \circ p_1\\
 	d\circ \pi_1\circ p_2&=d\circ n'\circ \pi^1_b=n\circ b \circ \pi^1_b=n\circ b\circ \pi^2_b=d\circ n'\circ \pi^2_b=d\circ \pi_2 \circ p_2
 \end{align*}
	so that $d\circ \pi_1=d\circ p_2$.

	We want now to show that $(P, \pi_1, \pi_2)$ is a kernel pair for $d$. Let $aa$

We 
For every $t$ in $T$
